---
title: "	Inside-Out Statistics"
subtitle: 'Teaching Evidence-Based Reasoning in Introductory Courses'
author: "Kelly Bodwin"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      countIncrementalSlides: no
      highlightLines: yes
      highlightStyle: github
institute: 'These slides: **kellybodwin.com/talks/jsm2020**'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(flair)
#library(icon)
#library(emo)
library(tidyverse)
library(knitr)

```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_light(
  base_color = "#26116c",
  text_bold_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  background_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```

```{css, echo = FALSE}
img {
  height: 400px;
}

.remark-code {
  font-size: 30px;
}

.remark-slide {
  background-size: cover;
}
```

---
class: center, middle

# A quick grading exercise

---

# The Question:

In 1952, a group of scientists decided to study the effect of Vitamin C supplements
on tooth growth.  They gathered a set of 60 guinea pigs, and gave half the animals
pure vitamin C and half the animals Orange Juice.  They then measured the length
of the guinea pigs' incisor teeth.

The results of the study are summarized below:

```{r, echo = FALSE}
ToothGrowth %>%
  group_by(supp) %>%
  summarize(`Mean Tooth Length` = mean(len), 
            `SD of Tooth Length` = sd(len),
            `Sample Size` = n()) %>%
  mutate(
    `Supplement Type` = factor(supp, 
                               levels = c("OJ", "VC"), 
                               labels = c("Orange Juice", "Vitamin C")),
  ) %>%
  select(-supp) %>%
  kable()
```

Is there evidence at the 0.05 level that the orange juice is more beneficial to tooth
growth than Vitamin C supplements?  Justify your answer.

---
# The Correct Answer

```{r, include = FALSE}
t.test(len ~ supp, ToothGrowth, alternative = "greater")
```

<p style="border:3px; border-style:solid; border-color:#6495ed; padding: 1em;">
With a t-score of <b>1.91</b>, we get a p-value of <b>0.03</b>.

There <b>is evidence</b> at the 0.05 level that Orange Juice leads to more tooth
growth than Vitamin C.
</p>

---
# Student A

<p style="border:3px; border-style:solid; border-color:red; padding: 1em;">
With a t-score of <b>1.91</b>, we get a p-value of <b>0.03</b>.  

There <font color = "red"><b> is not evidence</b></font> at the 0.05 level that Orange 
Juice leads to more tooth
growth than Vitamin C.
</p>

--

<p style="border:3px; border-style:solid; border-color:#6495ed; padding: 1em;">
With a t-score of <b>1.91</b>, we get a p-value of <b>0.03</b>.  

There <b>is evidence</b> at the 0.05 level that Orange Juice leads to more tooth growth than Vitamin C.
</p>

---
# Student B

<p style="border:3px; border-style:solid; border-color:red; padding: 1em;">
With a t-score of <font color = "red"><b>-1.91</b></font>, we get a p-value of <font color = "red"><b>0.03</b></font>.  

There <b>is evidence</b> at the 0.05 level that <font color = "red">Vitamin C</font> leads to more tooth growth than <font color = "red">Orange Juice</font>.
</p>

<p style="border:3px; border-style:solid; border-color:#6495ed; padding: 1em;">
With a t-score of <b>1.91</b>, we get a p-value of <b>0.03</b>.  

There <b>is evidence</b> at the 0.05 level that Orange Juice leads to more tooth
growth than Vitamin C.
</p>

---
# Student C


<p style="border:3px; border-style:solid; border-color:red; padding: 1em;">
<font color = "red">The standard deviation of the mean for the OJ group is 1.2.
The mean of the OJ group is about 4mm above the mean of the VC group.  This
difference is a lot bigger than the standard deviation. </font>

Thus, I think there <b>is evidence</b> that Orange Juice leads to more tooth
growth than Vitamin C.
</p>


--

<p style="border:3px; border-style:solid; border-color:#6495ed; padding: 1em;">
With a t-score of <b>1.91</b>, we get a p-value of <b>0.03</b>. 

There is <b>is evidence</b> at the 0.05 level of a difference in mean tooth lengths
across the two supplement types.
</p>

---
class: center, middle

# How did you score these students?

---

# Who got what wrong?

**Student A** got
* the **correct** t-score
* the **correct** p-value
* the **wrong** conclusion

**Student B** got
* the **correct** (ish) t-score
* the **correct** p-value
* the **wrong** conclusion

**Student C** got
* **no** t-score
* **no** p-value
* the **correct** conclusion


---
class: center, middle

# Which student would you want to hire?

---


**Student A:**


<p style="border:3px; border-style:solid; padding: 1em;">
With a t-score of <b>1.91</b>, we get a p-value of <b>0.03</b>.  

There <b> is not evidence</b> at the 0.05 level that Orange 
Juice leads to more tooth
growth than Vitamin C.
</p>

**Student B:**

<p style="border:3px; border-style:solid; padding: 1em;">
With a t-score of <b>-1.91</b>, we get a p-value of <b>0.03</b>.  

There <b>is evidence</b> at the 0.05 level that Vitamin C leads to more tooth growth than Orange Juice.
</p>

**Student C:**

<p style="border:3px; border-style:solid; padding: 1em;">
The standard deviation of the mean for the OJ group is 1.2.
The mean of the OJ group is about 4mm above the mean of the VC group.  This
difference is a lot bigger than the standard deviation. 

Thus, I think there <b>is evidence</b> that Orange Juice leads to more tooth
growth than Vitamin C.
</p>

--

I argue that **Student C** deserves the highest grade.

---
class: center, middle

# Being able to draw 
# **reasonable conclusions from data**
# is *way more important* than
# being able to perform a hypothesis test.

---

> It will not be
helpful for students to know about the tools and procedures that can be used to analyze data if
students don’t first understand the underlying concepts. Having a good understanding of the
concepts will make it easier for students to use necessary tools and procedures to answer
particular questions about a dataset.
Procedural steps too often claim students’ attention that an effective teacher could otherwise
direct toward concepts.  

> - Guidelines for Assessment and Instruction in Statistics Education (GAISE), 2016

---
# A Traditional Statistics Class

[pic of "mountain climbing" through the steps of summaries & viz, test stats,
p-values, interpretation]


---
# An "Inside-Out" Statistics Class

[pic of target with "reasoning from summaries", then "reasoning from test stats"
then "formal testing"]

---
# 
